---
title: "Homework 2"
author: Yue Liu
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

**Read the Mr. Trashwheel dataset**

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )      

```

**Read preciptation data for 2018 and 2017**

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

**Combine annual precipitation**

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr.Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it iin a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total `r nrow(trashwheel_df)` rows in our final dataset. From the dataset, we can know that there are `r filter(trashwheel_df, year == 2017) %>% select(sports_balls) %>% sum()` sports balls among the total `r filter(trashwheel_df, year == 2017) %>% select(weight_tons) %>% sum()` tons of trash being collected in 2017.

Additional data sheets include month precipitation data. According to the data, the total precipitation in 2017 was `r select(precip_2017, col = 3) %>% sum()`, and the total precipitatioh in 2018 was also `r select(precip_2018, col = 3) %>% sum()`. 

## Problem 2

**Read the NYC Transit Entrance and Exit dataset**

```{r}
transit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, contains("route"), entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, 'YES' = TRUE, 'NO' = FALSE)) %>% 
  mutate(route8 = as.factor(route8)) %>% 
  mutate(route9 = as.factor(route9)) %>% 
  mutate(route10 = as.factor(route10)) %>% 
  mutate(route11 = as.factor(route11))
```

The current dataset contains `r ncol(transit_df)` variables: line, station, name, station latitude, station longitude, routes served (1 - 11), entry, vending, entrance type, and ADA compliance. I have cleaned the variable names to let the names only consist "_", numbers, and letters. I have also converted the entry variable from character to a logical variable and the route variables to factor variables. The dimension for this dataset is `r nrow(transit_df)` rowa x `r ncol(transit_df)` columns. I don't think these data are tidy enough, since I find the dataset is hard to read with a lot of NA values present in the route columns. 

There are `r distinct(transit_df) %>% nrow()` distinct stations.
`r filter(transit_df,ada == TRUE) %>% nrow()`stations are ADA compliant.
There are `r filter(transit_df,vending =="NO") %>% nrow()` station entrances/exits without vending allow entrance, which is `r 183/1868 * 100`% of the total stations. 

**Create distinct variables: route number and route name**

```{r}
transit_df_tidy = 
  transit_df %>% 
    pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name") %>% 
  drop_na()
```

There are `r filter(transit_df_tidy, route_name =="A") %>% distinct() %>% nrow()` distinct stations serve the A train.
Among the distinct stations that serve A train, `r filter(transit_df_tidy, route_name =="A", ada == TRUE) %>% distinct() %>% nrow()` stations are ADA compliant, and among all the stations that serve A train, `r filter(transit_df_tidy, route_name =="A", ada == TRUE) %>% nrow()` stations are ADA compliant.


