---
title: "Homework 2"
author: Yue Liu
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

**Read the Mr. Trashwheel dataset**

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )      

```

**Read preciptation data for 2018 and 2017**

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

**Combine annual precipitation**

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr.Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total `r nrow(trashwheel_df)` rows in our final dataset. From the dataset, we can know that there are `r filter(trashwheel_df, year == 2017) %>% select(sports_balls) %>% sum()` sports balls among the total `r filter(trashwheel_df, year == 2017) %>% select(weight_tons) %>% sum()` tons of trash being collected in 2017.

Additional data sheets include month precipitation data. According to the data, the total precipitation in 2017 was `r select(precip_2017, col = 3) %>% sum()`, and the total precipitatioh in 2018 was also `r select(precip_2018, col = 3) %>% sum()`. 

## Problem 2

**Read the NYC Transit Entrance and Exit dataset**

```{r}
transit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, contains("route"), entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, 'YES' = TRUE, 'NO' = FALSE)) %>% 
  mutate(route8 = as.factor(route8)) %>% 
  mutate(route9 = as.factor(route9)) %>% 
  mutate(route10 = as.factor(route10)) %>% 
  mutate(route11 = as.factor(route11))
```

The current dataset contains `r ncol(transit_df)` variables: line, station, name, station latitude, station longitude, routes served (1 - 11), entry, vending, entrance type, and ADA compliance. I have cleaned the variable names to let the names only consist "_", numbers, and letters. I have also converted the entry variable from character to a logical variable and the route variables to factor variables. The dimension for this dataset is `r nrow(transit_df)` rows x `r ncol(transit_df)` columns. I don't think these data are tidy enough, since I find the dataset is hard to read with a lot of NA values present in the route columns. 

There are `r distinct(transit_df) %>% nrow()` distinct stations.
`r filter(transit_df,ada == TRUE) %>% nrow()` stations are ADA compliant.
There are `r filter(transit_df,vending =="NO") %>% filter(entry == TRUE) %>% nrow()` station entrances/exits without vending allow entrance and `r filter(transit_df,vending =="NO") %>% nrow()` station entrances / exits without vending. Therefore the proportion of station entrances/exits allow entrance among all the station entrances/exits that do not hace vending is `r 69 / 183 * 100` %. 

**Create distinct variables: route number and route name**

```{r}
transit_df_tidy = 
  transit_df %>% 
    pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name") %>% 
  drop_na()
```

There are `r filter(transit_df_tidy, route_name =="A") %>% distinct() %>% nrow()` distinct stations serve the A train.
Among the distinct stations that serve A train, `r filter(transit_df_tidy, route_name =="A", ada == TRUE) %>% distinct() %>% nrow()` stations are ADA compliant, and among all the stations that serve A train, `r filter(transit_df_tidy, route_name =="A", ada == TRUE) %>% nrow()` stations are ADA compliant.


## Problem 3

**Read the pols_month dataset**

```{r}
pols_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, c("year","month","date"), "-") %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.name[month]) %>% 
  select(-date) %>% 

  pivot_longer(
    starts_with("prez_"),
    names_to = "president",
    names_prefix = "prez_",
    values_to = "indicator") %>% 
  
  mutate(indicator = as.factor(indicator)) %>% 
  mutate(indicator = recode(indicator, '1' = "yes", '0' = "no"))
```

The "pols_month" dataset contains information on the numbers of politicians from different democratic or republican positions on any given months, ranging from January 1947 to June 2015. The dataset contains `r nrow(pols_df)` rows x `r ncol(pols_df)` columns. It has ten variables: the year of the count, the month of the count, numbers of democrats from three different positions, numbers of republicans from three different positions and indicators of whether the president is democratic or republican.

**Read the snp dataset**

```{r}
snp_df = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, c("month","date","year"), "/") %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.name[month]) %>% 
  select(-date) %>% 
  relocate(year, month)
```

The "snp" datasets contains information on the Standard & Poor’s stock market index, which is used as a representative measure of the stock market, on any given months, ranging from January 1950 to July 2015. It has `r nrow(snp_df)` rows x `r ncol(snp_df)` columns, containing 3 different variables, which are the observed year, month, and the S&P index during the associated time.

**Read the unemployment dataset**

```{r}
unemployment_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 

  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_percentage")
```
  
```{r}
month_df_2 = 
  tibble(
    month =  c("jan","feb","mar","apr","may","jun","jul","aug","sep","oct","nov","dec"),
    month_name = month.name
  )

unemployment_df_tidy = (
  left_join(unemployment_df, month_df_2) %>% 
    select(-month) %>%
    rename(month = month_name) %>% 
    mutate(year = as.factor(year)) %>% 
    drop_na() %>% 
    relocate(year, month, unemployment_percentage))
```

The “unemployment” dataset contains information on the unemployment percentage in any given month, ranging from January 1948 to June 2015. It has `r nrow(unemployment_df_tidy)` rows x `r ncol(unemployment_df_tidy)` columns, containing 3 different variables, which are the observed year, month, and the unemployment percentage during the associated time. 

**Join the datasets**

```{r}
ps_df = 
  left_join(pols_df, snp_df) 
```

```{r}
result_df = 
  left_join(ps_df, unemployment_df_tidy)
```

The final resulting dataset contains all the information present in the "pols_month" dataset, the "snp" dataset, and the "unemployment" dataset from January 1947 to June 2015. It has `r nrow(result_df)` rows x `r ncol(result_df)` columns, containing 12 different variables.